Cross-lingual Document Similarity and Event Detection \todo{Better title...}

### 2015-02-21 rupnik2015.tex

\section{Multilingual Latent Factor Models}

>>>>

\section{Multilingual Latent Factor Models}\label{sec:models}

\subsection{Multilingual dataset}

The cross-lingual similarity models presented in this paper are based on comparable corpora. That is, a corpus of documents in multiple languages, with alignment between documents that are of the same topic, or even a rough translation of each other. Wikipedia is an example of a comparable corpus, where a specific entry can be described in multiple languages (e.g. "Berlin" is currently described in 222 languages). News articles represent another example, where the same event can be described by newspapers in several languages.

More formally, \emph{multilingual document} $d = (u_1,\ldots u_m)$ is a tuple of $m$ documents on the same topic (comparable), where $u_i$ is the document written in language $i$. Note that individual document $u_i$ can be an empty document (missing resource) and each $d$ must contain at least two nonempty documents. A comparable corpus $D = {d_1, \ldots, d_N}$ is a collection of multilingual documents. By using the vector space model we can represent $D$ as a set of $m$ matrices $X_1,\ldots,X_m$, where $X_i \in \RR^{n_i \times N}$ is the matrix corresponding to the language $i$ and $n_i$ is the vocabulary size of language $i$. Furthermore, let $X_i^{\ell}$ denote the $\ell$-th column of matrix $X_i$ and the matrices respect the document alignment - the vector $X_i^\ell$ corresponds to the TFIDF vector of the $i$-th component of multilingual document $d_\ell$. We use $N$ to denote the total row dimension of $X$, i.e. $N:= \sum_{i=1}^m n_i$.
